paper1_figures

### 2023-12-18 scoping_review_short.Rmd
18.12 (=> inclusion of feedback)

abstract: |
  Commons represent an alternative to the classical capitalist model, therefore they could be the needed change toward a more sustainable society and economy. However, for a commons to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in the scientific field of the commons. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since social dilemma experiments and commons share some structural characteristics, we could gain insights into cooperation in commons and how to implement it thanks to the social dilemmas literature.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, in order to better understand what could shape cooperation in commons. A key-word search and a subsequent selection led to the inclusion of 89 papers, which were used to identify 9 factors that can influence cooperation both at the individual and group level: gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity. These factors are discussed within the existing literature on commons and cooperation. More specifically, the influences that these elements can have on each-other are highlighted, showing that all factors should be taken into account when investigating cooperation in commons. Finally, the practical implication of this work for commons' organization, as well as the review's limitations, are pointed out.

@@

floatsintext      : no
linenumbers       : yes
draft             : no
mask              : no

@@

In recent years, the interest in alternative ways of social organization increased drastically due to the social and ecological crises we are facing [@Euler2018]. Commons are one of these alternatives, since they are a way of organizing resources beyond the classic private and public model [@Ostrom1990], and, therefore, they represent an option to the capitalistic system [@Euler2019]. For this reason, commons could be an integral part of the path to attain a more sustainable and fair social organization [@Schmelzer2022].

Commons have been defined in different ways, starting by Elinor Ostrom who, in her seminal work, refers to them as a resource system shared between a self-managed community of users [@Ostrom1990]. This community shares a set of social practices, values and norms. Therefore, it is important to consider the commons not only as a resource, but as the integration of the resource, the community and its social protocols [@Bollier2014]. For this reason, beyond the material aspect, the notion of social practices (i.e. commoning) is essential [@Euler2018]. "Commoning" refers to the process through which people (i.e. the commoners) find solutions for their needs by self-organizing independently from the state or the market, and by creating autonomously managed institutions to govern the shared resource [@Euler2018; @Schmelzer2022]. The community is therefore engaged in defining rules to successfully manage the commons [@Meroni2022]. In conclusion, commons can be understood as "living social systems" that allow commoners to self-organize to address shared problems [@Bollier2019].

These definitions highlight the importance of the community as a part of the commons, and that is why it is essential to have a closer look at the people who form it. More specifically, in order for a commons[^1] to be successful, its members have to cooperate, i.e to incur a personal sacrifice for the greater good [@Kollock1998]. Although it is essential, cooperation can be difficult to achieve due to two characteristics of the commons' social dynamic. First, people can believe that their individual contribution will not have a big impact, and this feeling of personal inefficacy discourages cooperation [@VanVugt1998]. Secondly, in a commons, every member faces a conflict between individual, short-term interests and collective, long-term ones [@Hardin1968]. Act for the common good requires a personal sacrifice, and therefore the temptation to pursue the individual interest is constantly present. For these reasons, we need to understand the factors that affect cooperation, and how we can improve it.

@@

In the existing literature, there are several explications of why people adopt or not cooperative behaviours in commons. For instance, the Rational Choice Theory predicts that people are too self-interested to cooperate in such contexts. As a consequence, they will inevitably destroy the shared resource, unless an external institution forces them to cooperate [@DeCaro2021; @Hardin1968]. However, there are some less pessimistic theories that predict that people can transcend their personal interest and successfully cooperate in commons. An example is the Greed Efficiency Fairness hypothesis, which posits that the individuals' greed to pursue their narrow interests is constrained by their desire to use the resource efficiently and to fairly allocate it between members [@Wilke1991]. Similarly, according to the Humanistic Rational Choice Theory, the desire to maintain basic social-psychological needs motivates people to cooperate [@DeCaro2021]. Finally, cooperating in commons can also be a self-presentation strategy [@VanVugt2010a], since showing altruistic behaviours has a positive effect on reputation [@Roberts1998].

On a more practical level, @Ostrom1990 reported several cases of successful commons, where individuals managed to create a system to ensure cooperation. Based on her observations, she established eight design principles that distinguish successful from unsuccessful commons. First, commons need clearly defined boundaries to determine who has access to the shared resource, and the governing rules must take into account the local conditions. Moreover, people affected by these rules must be able to participate in modifying them. A monitoring and a graduated sanctions system must also be implemented, as well as low-cost conflict resolution arena. Finally, external authorities must not challenge the commons' members' right to self-organize[^2] [@Ostrom1990].

[^2]: The last principle is valid only for commons that are part of a larger system, and is about nested enterprises: all the governance activities should be organized in multiple layers of nested enterprises.

There are therefore several studies that focused on commons and on how they can be successful, but social psychological and quantitative perspectives are missing from this literature. In fact, to the authors' knowledge, there are few studies that investigated the causal effect of individual and group factors on people's willingness to cooperate in a real-life commons [e.g. @Noonan2016].

At the same time, an experimental approach is very popular in the laboratory study of social dilemmas, which identified several factors that shape cooperation [e.g. @VanLange2013]. In these experiments, people face a conflict between personal and collective interest: each individual obtains a higher personal payoff with a non-cooperative, self-interested behaviour than with a social, cooperative one, independently from what others do. At the same time, everybody's payoff is higher if all cooperate than if all defect [@Dawes1980]. Since many of the current global problems have a social dilemma structure, this topic has been extensively studied in different fields such as psychology, economy, biology, or mathematics [@VanLange2013]. Social dilemmas are very useful to gain insights into how cooperation works, because they can easily be implemented in a laboratory setting, and they provide a standard experimental paradigm to test under which conditions people cooperate [@Balliet2011].

@@

Cooperation can be modified by introducing incentive systems, that can be either punishments or rewards. Both have a positive effect on cooperation [@Balliet2011a]. However, some studies found that punishments are more efficient than rewards [e.g. @Milinski], while others came to the opposite conclusion [e.g. @Chen2022].

>>>>

abstract: |
  Private ownership and commodification of resources are central aspects of the current capitalistic social organization. This has detrimental effects both on social and environmental levels. Commons represent an alternative to this system, therefore they could be the needed change toward a more sustainable society and economy. However, for a commons to be successful, its members must cooperate, i.e. incur a personal sacrifice for the collective good. Despite the importance of cooperation, there is a lack of studies that investigated this topic in the scientific field of the commons. On the other hand, cooperation has been extensively studied in social dilemma experiments, where personal and collective interests are at odds. Since social dilemma experiments and commons share some structural characteristics, we could gain insights into cooperation in commons and how to promote it thanks to the social dilemmas literature.
  The goal of the current scoping review is to identify the factors that impact cooperation in social dilemma situations, in order to better understand what could shape cooperation in commons. A key-word search and a subsequent selection led to the inclusion of 89 papers, which were used to identify 9 factors that influence cooperation both at the individual and group level: gender, social status, group identification, values and personality traits, trust, incentives, communication, social norms and anonymity. The influences that these elements can have on each-other are highlighted, showing that all factors should be taken into account when investigating cooperation in commons. Finally, the theoretical and practical implications of this work for commons' organization, as well as the review's limitations, are pointed out.

@@

floatsintext            : no
linenumbers             : yes
draft                   : no
mask                    : no
disambiguate_authors    : no

@@

<!-- change of the initial paragraph and of the definition -->

In recent years, the interest in alternative ways of social organization increased drastically due to the social and ecological crises we are facing [@Euler2018]. Commons could be an integral part of the path to attain a more sustainable and fair social organization [@Schmelzer2022]. In fact, they are a way of organizing resources beyond the classic private and public model [@Ostrom1990] and they are not organized around market competition [@Euler2019]. Commons express "a broader conception of property, referring to social goods [...] which a community, not the state or any individual, collectively owns, manages, and controls" [@Federici2011, p. 41]. Giving their characteristics, commons represent an option to the capitalistic system [@Euler2019] and its detrimental effects on natural resources and human well-being.

Commons can assume various forms, because "they can arise whenever a community decides it wishes to manage a resource in a collective manner, with a special regard for equitable access, use and sustainability" [@Bollier2014, p.19]. The variety of definitions can be resumed through three key points. First, commons allow to manage shared resources [@Ostrom1990] and to find solution for share needs of the population [@Euler2018; @Euler2019]. Second, commons are managed by a self-organized community of users [@Ostrom1990], i.e. the commoners. This community is a "social network of mutual aid, solidarity, and practices of human exchange that are not reduced to the market form" [@DeAngelis2003, p.1]. Moreover, commoners' organization is usually local, egalitarian, integrative, and fair [@Gidwani2011]. Third, the community's social practices are called "commoning" and, thanks to this process, commoners create institutions to govern the commons [@Euler2018; @Schmelzer2022]. Finally, although commons are often thought just as a shared resource, the other two aspects are also essential in defining them. Therefore, it is important to consider the commons as the integration of the resource, the community (i.e. the commoners) and its social practices (i.e. commoning) [@Bollier2014]. These definitions highlight the importance of the community as a part of the commons, and that is why it is essential to have a closer look at the people who form it.

More specifically, in order for a commons[^1] to be successful, its members have to cooperate, i.e to incur a personal sacrifice for the greater good [@Kollock1998]. Although it is essential, cooperation can be difficult to achieve due to two characteristics of the commons' social dynamic. First, people can believe that their individual contribution will not have a big impact, and this feeling of personal inefficacy discourages cooperation [@VanVugt1998]. Secondly, in a commons, every member faces a conflict between individual, short-term interests and collective, long-term ones [@Hardin1968]. Since acting for the common good requires a personal sacrifice, the temptation to pursue the individual interest is constantly present. Giving the role that commons could play in addressing the environmental and social problems, we need to understand the factors that affect cooperation, and how we can improve it. In fact, despite its importance, the reasons why individual choose to cooperate have not been extensively studied in commons scientific literature. At the same time, social dilemma experiments (i.e. artificial decision making situations, where personal and collective interests are at odds) allowed to identify several factors that could impact cooperation [e.g. @VanLange2013]. Since commons and social dilemmas share some structural characteristics [@Hardin1968; @Dawes1980], the aim of the current scoping review is to gain insight into cooperation in commons through social dilemma literature.

@@

<!-- the whole structure has been changed according to Lilla's and Steffi's suggestions + Ostrom's principles explained more in details -->

One of the first paper about cooperation in commons was Hardin's "The tragedy of the commons" [-@Hardin1968]. In this article, Hardin argues that commons' members are rational beings who seek to maximize their personal interest by over-exploiting the common resource. This would inevitably lead to the destruction of the resource, unless an external institution regulates the access, or the resource is privatized. Therefore, according to Hardin, people benefiting from a common resource cannot successfully manage it without external interventions.

However, some years later, Elinor Ostrom's seminal work [-@Ostrom1990] falsified Hardin's predictions about people's inability to manage a common resource and their greed to maximize their self-interest. In fact, Ostrom observed several real-life commons, and showed that some of them were very successfully managed by the community of users without the intervention of an external authority. Based on her observations, she established seven design principles that are present in all the observed successfull commons, plus an eight one for more complex cases.[^2] By "design principle", Ostrom means "an essential element or condition that helps to account for the success of these institutions" [@Ostrom1990, p. 90].

[^2]: Since the eight principle is not relevant for the current paper, it will not be detailed.

The first principle posits that clearly defined boundaries are needed: it must be clear which is the shared resource, and who can benefit from it, otherwise, "no one knows what is being managed or for whom" (p. 91). Second, the rules must be congruent with the local conditions, i.e. they must reflect the specific attributs of the particular resource. Thirdly, individuals affected by the operational rules must be able to participate in modifying them. This ensure that the rules fit well to the local conditions and allows to promptly adapt them when needed. The fourth and fifth principles concern a monitoring and a graduate sanction system. Interestingly, these systems are successfully implemented by commoners them-selves, that can assume the roles of monitor or punisher, with no need of an external institution. In fact,

-   everybody has a self-interest in monitoring the others

-   no one will abuse of the sanctions because they know that they could be in the same situation so their are comprehensive + being caught has already a negative effect on reputation

1.  Monitoring + graduated sanctions (4 and 5) =\> these two systems are implemented by commoners them-selves, not by external institutions
    1.  Monitoring = low cost, because it is in everybody's self interest to monitor others + being a good monitor and catch somebody's infraction can make a person gain in prestige and status + material benefits if an infraction is caught (ex: part of the fine) + monitoring necessary to adopt a contingent strategy (if I cooperate, I want to be sure that the others do not exploit my work)
    2.  Not enourmous sanctions needed, because being caught is already negative for reputation
2.  Conflict resolution mechanism (6) =\> individuals following rules for a long period of time = inevitably some infractions, so it is essential to have somewhere to discuss when conflicts arise =\> give the possibility to explain the mistake, so that compliace to rules does not drop
3.  Minimal recognition of the right to organize (7) =\> external authorities (such governement) should not challenge the commoners' right to organize (if they do it, the rules of appropriators are not seen as valid and therefore more difficultly compliable)

These principles were not present in Hardin's example, where the hypothetical commoners made decisions in a context without local organization or established norms. Therefore, Hardin's scenario should not be considered a commons, but an open access regime [@Ostrom2008; @Bollier2014], and this can contribute to explain his negative predictions about cooperation.

At the same time, psychology is so far absent from the discussion

-   psychology is absent so far in the discussion, even though there are theories and experimental paradigms that are highliy relevant to the topic. On the theory side, there is....[the first paragraph with the different Theories] On the experimental side, there are classical social psychological experiments that...

Secondly, Hardin assumed, in line with the Rational Choice Theory, that all the commoners wanted to maximize their personal interest [@Wittel2023]. However, there are several theories less pessimistic than the RCT about human nature. For instance, the Greed Efficiency Fairness hypothesis posits that the individuals' greed to pursue their narrow interests is constrained by their desire to use the resource efficiently and to fairly allocate it between members [@Wilke1991]. Similarly, according to the Humanistic Rational Choice Theory, the desire to maintain basic social-psychological needs motivates people to cooperate [@DeCaro2021]. Moreover, cooperating can also be a self-presentation strategy [@VanVugt2010a], since showing altruistic behaviours has a positive effect on reputation [@Roberts1998].

Finally, the laboratory study of social dilemmas also largely contributed to the investigation of this topic, by allowing researchers to identify several factors that shape cooperation [e.g. @VanLange2013]. In general, these studies falsified RCT by demonstrating that, in social dilemmas, people cooperated more than predicted, and that their behaviour was impacted by variables that should not play a role according to the theory [@Ostrom1998]. Moreover, research based on different fields has convincingly demonstrated that people can engage in collective actions to sustainably manage shared resources [@Poteete2010a].

CONCLUSION: PSYCHOLOGY OFFER A SECOND EXPLANAITION OF WHY THERE IS NOT ALWAYS A TRAGEDY OF THE COMMONS = PEOPLE DO NOT ONLY WANT TO MAXIMISE THEIR PERSONAL INTEREST, AND SEVERAL "IRRATIONAL" FACTORS IMPACT THE WAY IN WHICH THEY INTERACT WITH OTHERS

@@

Cooperation can be modified by introducing incentive systems, that can be either punishments or rewards. Both have a positive effect on cooperation [@Balliet2011a]. However, some studies found that punishments are more efficient than rewards [e.g. @Milinski2012], while others came to the opposite conclusion [e.g. @Chen2022].
